---
title: "Analysis Master"
author: "Ethan Walker"
date: "November 1, 2017"
output: html_document
---

```{r, include=FALSE}
library(readxl)
library(tidyverse)
library(ggrepel)
library(gghighlight)
```

#Association between dependent and independent variables:
Create summaries and visualizations of how the dependent variable is
associated with different independent variables. Here, we will try to
discover if there are characteristics of the drugs that are associated
with effectiveness against TB. This group will need to come up with
ways (and code) to analyze that in the data. This might include
generalized linear models, scatterplots, and possibly other supervised
learning methods.

#Pull in Data
```{r}
efficacy_summary <- read_csv("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/master/CSV_Files/efficacy_summary.csv")
head(efficacy_summary)

clean_2_combined <- read_csv("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/master/CSV_Files/clean_2_combined.csv")
head(clean_2_combined)
```