---
title: "Ethan Notes and Examples"
author: "Ethan Walker"
date: "November 11, 2017"
output: html_document
---
```{r, include=FALSE}
library(readxl)
library(tidyverse)
library(ggrepel)
library(gghighlight)
library(ggplot2)
library(rpart)
library(party)
library(randomForest)
library(ggthemes)
library(rattle)
library(rpart.plot)
library(RColorBrewer)
source("practice_data.R")
```

#Association between dependent and independent variables:
Create summaries and visualizations of how the dependent variable is
associated with different independent variables. Here, we will try to
discover if there are characteristics of the drugs that are associated
with effectiveness against TB. This group will need to come up with
ways (and code) to analyze that in the data. This might include
generalized linear models, scatterplots, and possibly other supervised
learning methods.

Next:
Regression Tree - Ethan (also work on scatterplots more)
Random forest - Drew
Stepwise - Maggie
Univariate models for each independent var - Kate
GLM model - Kate

```{r}
efficacy_summary <- read_csv("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/master/CSV_Files/efficacy_summary.csv")
head(efficacy_summary)

clean_2_combined <- read_csv("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/master/CSV_Files/clean_2_combined.csv")
head(clean_2_combined)
```


```{r}
cmax <- read_excel("../data/DRUG_CLASS_I_Mean_Cmax_Trough_Efficacy_R_DATA_ANALYSIS.xlsx", 
                   sheet = "Mean_PK_Efficacy_In vitro", skip = 1, n_max = 12)

trough <- read_excel("../data/DRUG_CLASS_I_Mean_Cmax_Trough_Efficacy_R_DATA_ANALYSIS.xlsx", 
                     sheet = "Mean_PK_Efficacy_In vitro", skip = 16, n_max = 12)
```


Start cleaning data up
```{r}
cmax_clean <- cmax %>%
  rename(drug = X__1, dose = Dose) %>% 
  gather(key = dependent, value = value_dependent, ELU:ESP) %>% 
  gather(key = independent, value = value_independent, -dependent, -value_dependent, - drug, -dose) %>% 
  mutate(value_dependent = as.numeric(value_dependent), value_independent = as.numeric(value_independent)) 
head(cmax_clean)
```


Example from ggrepel vignette
geom_text  
https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html
```{r}
library(ggplot2)
ggplot(mtcars) +
  geom_point(aes(wt, mpg), color = 'red') +
  geom_text(aes(wt, mpg, label = rownames(mtcars))) +
  theme_classic(base_size = 16)
```

geom_text_repel
```{r}
library(ggrepel)
set.seed(42)
ggplot(mtcars) +
  geom_point(aes(wt, mpg), color = 'red') +
  geom_text_repel(aes(wt, mpg, label = rownames(mtcars))) +
  theme_classic(base_size = 16)
```

geom_label_repel
```{r}
set.seed(42)
ggplot(mtcars) +
  geom_point(aes(wt, mpg), size = 5, color = 'grey') +
  geom_label_repel(
    aes(wt, mpg, fill = factor(cyl), label = rownames(mtcars)),
    fontface = 'bold', color = 'white',
    box.padding = 0.35, point.padding = 0.5,
    segment.color = 'grey50'
  ) +
  theme_classic(base_size = 16)
```

Scatterplots with our data
```{r}
ggplot(efficacy_summary) +
  geom_point(aes(ELU, PLA), size = 5, color = 'grey') +
  geom_label_repel(
    aes(ELU, PLA, fill = dosage, label = drug),
    fontface = 'bold', color = 'white',
    box.padding = 0.35, point.padding = 0.5,
    segment.color = 'grey50'
  ) +
  theme_classic(base_size = 16)
```

Trying gghighlight:
This lets you filter based on values that you specify, then highlights the points that meet the requirements.
```{r}
gghighlight_point(data = efficacy_summary, aes(ELU, PLA), ELU > 2 & PLA > 10000, label_key = drug)
```


Faceting with independent variables
Making data long to facet with
```{r}
efficacy_summary_long <- efficacy_summary %>% 
  gather(key = dependent, value = value_dependent, ELU:ESP) %>% 
  gather(key = independent_bio, value = value_independent_bio, PLA:SLE) %>%
  gather(key = independent_chem, value = value_independent_chem, cLogP:MacUptake) %>%
  mutate(value_dependent = as.numeric(value_dependent), 
         value_independent_bio = as.numeric(value_independent_bio), 
         value_independent_chem = as.numeric(value_independent_chem)) 
head(efficacy_summary_long)
```
```{r}
plot_1 <- efficacy_summary_long %>% 
  ggplot() +
  geom_point(aes(x = value_independent_bio, y = value_dependent)) +
  facet_wrap(~ independent_bio, scales = "free_x") +
  theme_few()
plot_1
```

## rpart package
[rpart1](https://www.statmethods.net/advstats/cart.html)
[rpart2](https://cran.r-project.org/web/packages/rpart/rpart.pdf)
[rpart example](https://gormanalysis.com/decision-trees-in-r-using-rpart/)

Practice with rpart package:
```{r}
# grow tree 
fit <- rpart(Kyphosis ~ Age + Number + Start,
  	method="class", data=kyphosis)

printcp(fit) # display the results 
plotcp(fit) # visualize cross-validation results 
summary(fit) # detailed summary of splits

# plot tree 
plot(fit, uniform=TRUE, 
  	main="Classification Tree for Kyphosis")
text(fit, use.n=TRUE, all=TRUE, cex=.8)

# create attractive postscript plot of tree 
#post(fit, file = "tree.ps", 
#  	title = "Classification Tree for Kyphosis")
```

rpart and party with our data:
```{r}
## Remove missing outcome data
efficacy_summary_2 <- efficacy_summary %>% 
  filter(!is.na(ELU)) %>% 
  filter(!is.na(ESP))

## Trying rpart package:
# drug + dosage + level
# PLA + ULU + RIM + OCS + ICS + SLU + SLE
# cLogP + huPPB + muPPB + MIC_Erdman + MICserumErd + MIC_Rv + Caseum_binding + MacUptake
fit_1 <- rpart(ELU ~  drug + dosage + level + 
                 PLA + ULU + RIM + OCS + ICS + SLU + SLE + 
                 cLogP + huPPB + muPPB + MIC_Erdman + MICserumErd + MIC_Rv + 
                 Caseum_binding + MacUptake,
               data = efficacy_summary_2, 
               control = rpart.control(cp = -1, minsplit = 4, minbucket = 1))

printcp(fit_1) # display the results 
plotcp(fit_1) # visualize cross-validation results 
summary(fit_1) # detailed summary of splits

par(mfrow = c(1,1.5), xpd = NA) # otherwise on some devices the text is clipped
plot(fit_1, uniform=TRUE, 
  	main="Classification Tree for ELU")
text(fit_1, use.n=TRUE, all=TRUE, cex=.8)
#Make a fancy plot
fancyRpartPlot(fit_1)

## Trying party package:
fit_2 <- ctree(ELU ~  SLE + ICS + OCS + RIM + ULU + PLA + SLU, 
               data = efficacy_summary_2)

plot(fit_2, main="Conditional Inference Tree for ELU")

```




