---
title: "Ethan Notes and Examples"
author: "Ethan Walker"
date: "October 31, 2017"
output: html_document
---
```{r, include=FALSE}
library(readxl)
library(tidyverse)
library(ggrepel)
library(gghighlight)
library(ggplot2)
library(rpart)
library(randomForest)
source("practice_data.R")
```

#Association between dependent and independent variables:
Create summaries and visualizations of how the dependent variable is
associated with different independent variables. Here, we will try to
discover if there are characteristics of the drugs that are associated
with effectiveness against TB. This group will need to come up with
ways (and code) to analyze that in the data. This might include
generalized linear models, scatterplots, and possibly other supervised
learning methods.

Next:
Regression Tree - Ethan (also work on scatterplots more)
Random forest - Drew
Stepwise - Maggie
Univariate models for each independent var - Kate
GLM model - Kate

```{r}
efficacy_summary <- read_csv("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/master/CSV_Files/efficacy_summary.csv")
head(efficacy_summary)

clean_2_combined <- read_csv("https://raw.githubusercontent.com/KatieKey/input_output_shiny_group/master/CSV_Files/clean_2_combined.csv")
head(clean_2_combined)
```


```{r}
cmax <- read_excel("../data/DRUG_CLASS_I_Mean_Cmax_Trough_Efficacy_R_DATA_ANALYSIS.xlsx", 
                   sheet = "Mean_PK_Efficacy_In vitro", skip = 1, n_max = 12)

trough <- read_excel("../data/DRUG_CLASS_I_Mean_Cmax_Trough_Efficacy_R_DATA_ANALYSIS.xlsx", 
                     sheet = "Mean_PK_Efficacy_In vitro", skip = 16, n_max = 12)
```


Start cleaning data up
```{r}
cmax_clean <- cmax %>%
  rename(drug = X__1, dose = Dose) %>% 
  gather(key = dependent, value = value_dependent, ELU:ESP) %>% 
  gather(key = independent, value = value_independent, -dependent, -value_dependent, - drug, -dose) %>% 
  mutate(value_dependent = as.numeric(value_dependent), value_independent = as.numeric(value_independent)) 
head(cmax_clean)
```


#Example from ggrepel vignette
##geom_text  
https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html
```{r}
library(ggplot2)
ggplot(mtcars) +
  geom_point(aes(wt, mpg), color = 'red') +
  geom_text(aes(wt, mpg, label = rownames(mtcars))) +
  theme_classic(base_size = 16)
```

##geom_text_repel
```{r}
library(ggrepel)
set.seed(42)
ggplot(mtcars) +
  geom_point(aes(wt, mpg), color = 'red') +
  geom_text_repel(aes(wt, mpg, label = rownames(mtcars))) +
  theme_classic(base_size = 16)
```

##geom_label_repel
```{r}
set.seed(42)
ggplot(mtcars) +
  geom_point(aes(wt, mpg), size = 5, color = 'grey') +
  geom_label_repel(
    aes(wt, mpg, fill = factor(cyl), label = rownames(mtcars)),
    fontface = 'bold', color = 'white',
    box.padding = 0.35, point.padding = 0.5,
    segment.color = 'grey50'
  ) +
  theme_classic(base_size = 16)
```

#Scatterplots with our data

```{r}
ggplot(cmax_clean) +
  geom_point(aes(ELU, PLA), size = 5, color = 'grey') +
  geom_label_repel(
    aes(ELU, PLA, fill = Dose, label = drug),
    fontface = 'bold', color = 'white',
    box.padding = 0.35, point.padding = 0.5,
    segment.color = 'grey50'
  ) +
  theme_classic(base_size = 16)
```

##Trying gghighlight
  
This lets you filter based on values that you specify, then highlights the points that meet the requirements.
```{r}
gghighlight_point(data = cmax_clean, aes(ELU, PLA), ELU > 2 & PLA > 10000, label_key = drug)
```

```{r}
plot_1 <- cmax_clean %>% 
  ggplot() +
  geom_point(aes(x = value_independent, y = value_dependent)) +
  facet_wrap(~ independent, scales = "free_x")
plot_1
```



Trying plots with new cleaned data from other group:
```{r}

```


Practice with rpart package:
```{r}

```

