---
title: "Beeswarm practice"
author: "Kate Huebner"
date: "11/1/2017"
output: word_document
---

Modelling dependent~independent variables to determine variables associated with drug efficacy against TB. Dependent Variables are ELU (efficacy in the lungs) and ESP (efficacy in the spleen). Cmax and Trough indicate different drug concentrations and values are averaged across many mice at different drug doseages. Dependent variables include biologic and chemical (in vitro) measures. 

Here I will use beeswarm to plot scatterplots to show variation of different variables. ggbeeswarm provides two methods of plotting categorical scatter plots such that the arrangement of points within a category reflects the density of data at that region, and avoids over-plotting

```{r global_Options, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE, error = FALSE)
```

```{r echo=FALSE}
#Loading libraires
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(maps) 
library(readxl)
library(ggbeeswarm)

```

```{r}
#Reading in data
cmax <- read_excel("../data/DRUG_CLASS_I_Mean_Cmax_Trough_Efficacy_R_DATA_ANALYSIS.xlsx",
                    sheet = "Mean_PK_Efficacy_In vitro", skip = 1, n_max = 12)

trough <- read_excel("../data/DRUG_CLASS_I_Mean_Cmax_Trough_Efficacy_R_DATA_ANALYSIS.xlsx",
                   sheet = "Mean_PK_Efficacy_In vitro", skip = 16, n_max = 12)

head(cmax)
head(trough)

```
```{r}
#data cleaning
cmax <- cmax %>% 
  rename(Drug = X__1)
  

cmax
```

