---
title: "univariate models"
author: "Kate Huebner"
date: "11/1/2017"
output: word_document
---

Modelling dependent~independent variables to determine variables associated with drug efficacy against TB. Dependent Variables are ELU (efficacy in the lungs) and ESP (efficacy in the spleen). Sleen and lung are the main outcome variables. Cmax and Trough indicate different drug concentrations and values are averaged across many mice at different drug doseages. Dependent variables include biologic and chemical (in vitro) measures. 

Dependent variables:
ELU- concentration of drug in the lung
ESP- concentration of drug in the spleen

Cmax- Maximum volume of drug concentration at time x of treatment
Trough- lowest drug concentration after treatment


Kate assignment- Create a linear regression model fitwith lm. fit linear regression that are univariate and do separate variables by dose interval. 

```{r global_Options, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE, error = FALSE)
```

```{r echo=FALSE}
#Loading libraires
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(ggthemes)
library(readxl)
library(ggpmisc)
library(broom)
library(devtools)
library(ggfortify)

```

```{r}
#Reading in unclean data
source("practice_data.R")

```
```{r}
Cmax <- efficacy_summary %>% 
  filter(level == "Cmax") %>% 
  gather(key = independent_var, value = indep_measure, -drug, -dosage, -dose_int, -level, -ELU, -ESP, na.rm = TRUE) %>% 
  select(-ESP)
#add in rename of indep_variables here
head(Cmax)
```

```{r fig.width=8, fig.height=6}
#Scatterplot function- input dataframe, select what y is, cmax or trough, and then it returns the scatterplot
#this is set up for drug doseages are tied to the interval
scatter_plot <- Cmax %>% 
  ggplot(aes(x = indep_measure, y = ELU, color = dose_int)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Independent variable", y = "Dependent Variable") +
  facet_wrap(~independent_var, scales = "free_x")
  
scatter_plot
```


```{r}
#Univariate Models
#scale by independent measure: subtracting the mean value from the independent variable and dividing by the overall standard. that way the final mean is 0 and the st dv is 1. gives all variables a similar mean. that way the slope is comparable across everything
model_function <- function(df) {
    model_results <- lm(ELU ~ scale(indep_measure), data = df)
}
```

```{r}
#model function to output table
estimate_results <- Cmax %>% 
  group_by(independent_var, dose_int) %>% 
  nest() %>% 
  mutate(mod_results = purrr::map(data, model_function)) %>% 
  mutate(mod_coefs = purrr::map(mod_results, broom::tidy)) %>% 
  select(independent_var, dose_int, mod_results, mod_coefs) %>% 
  unnest(mod_coefs) %>% 
  filter(term == "scale(indep_measure)")
  
estimate_results
```
#use forcat:fct_reorder to graph scatterplot of coefficients on x and indep_var on y

```{r, fig.height = 4, fig.width = 7}
#we are reordering the indep var by the estimate value maximum. 
#bigger point size has less variance
coef_plot <- estimate_results %>%
  mutate(independent_var = forcats::fct_reorder(independent_var, estimate, fun = max)) %>%
  rename(Dose_Interval = dose_int) %>% 
  ggplot(aes(x = estimate, y = independent_var, color = Dose_Interval)) +
  geom_point(aes(size = 1 / std.error)) +
  scale_size_continuous(guide = FALSE) +
  theme_few() + 
  ggtitle(label = "Linear model coefficients as function of independent variables, \n by drug dose and model uncertainty", subtitle = "smaller points have more uncertainty than larger points") +
  geom_vline(xintercept = 0, color = "cornflower blue") 

coef_plot

```

#plot, the dependent variables (ELU  across individual mice). many mouse replicates for each drug. it was scaled so that they are all on the same scale so we can compare them
-trying to see if there are general trends across one drug class. is there a common feature that drives durg efficacy or activity. 
this model fit each feature what a time


```{r}
#individual mouse function
library(ggbeeswarm)
str(clean_2_combined)
head(clean_2_combined)

ind_mouse_data <- clean_2_combined %>% 
  select(drug, mouse_id, lung_efficacy) %>% 
  mutate(drug = as.factor(drug)) %>% 
  mutate(mouse_id = as.factor(mouse_id)) %>% 
  group_by(drug) 
  #Add mutate true or false logical statement if outside IQR
  
ind_mouse_plot <- ind_mouse_data %>%   
  ggplot(aes(x = drug, y = lung_efficacy)) +
  geom_beeswarm() +
  geom_smooth() +
  ylab("Colony Forming Units") +
  theme_few()
  
ind_mouse_plot
#Group by each drug and label for points outside IQR. then use Geomhighlight (and geomrepel (used in place of geomtext) to add on which mouse it was 
```

This plot shows the efficacy of each drug on in bacterial burden in the lung ( measured in CFU). Each data point is one mouse. There is a lot of variability in the mouse model relative to their pathology and infection
